---
title: "EPPS 6323 — Assignment 3"
author: "Poulami Sarker"
format: html
---

**In addition to the analysis presented in this document, the corresponding lab work (Lab_ML01) has been rendered and uploaded to the course website under the “Labs” section as Lab 3. The lab files include the full code, output, and visualizations used in the exploratory data analysis**

##3.Load Data:


```{r}
library(haven)

TEDS_2016 <- read_stata("https://github.com/datageneration/home/blob/master/DataProgramming/data/TEDS_2016.dta?raw=true")
```

## 4. Multiple Regression:

```{r}
model <- lm(Tondu ~ age + edu + income, data = TEDS_2016)
summary(model)
```

## Function: regplot:

```{r}
regplot=function(x,y){
 fit=lm(y~x)
 plot(x,y)
 abline(fit,col="red")
}
```

## 5. Regplot on the Dependent Variables


##Regression Plot: Age:
```{r}
regplot(TEDS_2016$age, TEDS_2016$Tondu)
```
##Regression Plot: Education:

```{r}
regplot(TEDS_2016$edu, TEDS_2016$Tondu)
```
##Regression Plot: Income:
```{r}
regplot(TEDS_2016$income, TEDS_2016$Tondu)
```
##6. What is the problem and why?

The dependent variable (Tondu) has multiple categories (more than two categories), so it is not a continuous variable. Linear regression assumes a continuous dependent variable with equal spacing between values, which is not appropriate here. Because Tondu is categorical with several categories, the regression model and plots are not suitable.

##7. How can prediction be improved?

To improve prediction, a model suitable for categorical outcomes should be used. For example, ordinal logistic regression or multinomial logistic regression would be more appropriate because they can handle multiple categories in the dependent variable without assuming equal distance between them.